---
title: "Final.project Report"
author: "Samira Bechi"
date: "2023-11-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## INTRODUCTION

[link for the Airbnb data](https://www.kaggle.com/datasets/PromptCloudHQ/airbnb-property-data-from-texas)
Founded in 2008, Airbnb is an online platform connecting hosts offering short-term lodging with travelers seeking alternatives to traditional hotels. Today, I'm analyzing Airbnb property data in Texas from 2008 to 2017, encompassing average night price, bedrooms, location, year listed, and descriptions. The analysis aims to address key questions about Texas listings, Airbnb's growth, popular cities, and average prices per night, per year, and per bedroom. The findings will be visually presented to offer insights into the Airbnb business.

## LIBRARY USED

```{r library}
library(kableExtra) # used to construct Complex Table for data
library(tidyverse) #using this to tidy up the data
library(dplyr) #using this for data manipulation
library(ggplot2) #using for visual data
```

## UNFILTERED DATA

```{r load_data}
load("Data.RData")
glimpse(Airbnb_Texas)
```

## FORMATING DATA

```{r format_data}
#removing all unecessary data
Airbnb_Texas$description <- NULL
Airbnb_Texas$latitude <- NULL
Airbnb_Texas$longitude <- NULL
Airbnb_Texas$url <- NULL
Airbnb_Texas$title <- NULL
# separate year and moth to be able to access data easily
date <- strsplit(x = as.character(Airbnb_Texas$date_of_listing) ,split = " ",fixed = T)
Airbnb_Texas$month <- as.factor(sapply(X = date,"[[",1))
Airbnb_Texas$year <- as.factor(sapply(X = date,"[[",2))
Airbnb_Texas$date_of_listing <- NULL
# remove $ sign and convert the average per night to a numeric 
Airbnb_Texas$average_rate_per_night <-substr(x = Airbnb_Texas$average_rate_per_night,start = 2,stop = 10)
Airbnb_Texas$average_rate_per_night <- as.numeric(Airbnb_Texas$average_rate_per_night)
# consider a studio as a 1 bedroom and change it to numeric 
Airbnb_Texas$bedrooms_count[Airbnb_Texas$bedrooms_count=="Studio"] <- "1"
Airbnb_Texas$bedrooms_count <- as.numeric(Airbnb_Texas$bedrooms_count)
glimpse(Airbnb_Texas)
```

## DATA KEY TERMS
**average_rate_per_night** cost of the room per night

**bedrooms_count** number of bedroom

**city** location of the airbnb

**date_of_listing** date when airbnb was available on platform

## 1- Total Airbnb from 2008 to 2017

From 2008 to 2017 exactly 18,259 were created in Texas.

## 2- Number of houses by year

We are taking our formatted data to display the number of houses per year. This will help us determine what year had the most additional Airbnbs and if there was an increase or decrease in the number of Airbnbs added yearly.

```{r}
ggplot(Airbnb_Texas %>% group_by(year) %>% count(), aes(year,n,fill=year,color=year)) +
   geom_col() +
   labs(title=" Number of houses by year in Texas",subtitle="By Year")
```

## 4- Top 10 city with most airbnb 

```{r}
Airbnb_Texas %>% group_by(city) %>% count() %>% arrange(desc(n)) %>% kbl() %>% row_spec(1:10, background = "pink") %>% kable_paper()  %>% scroll_box(height = "300px")
```

We can see that the top 10 cities have international airports and big companies, so the more traffic a town has, the more Airbnb there is.

## 5- Average rate per night per month 

```{r}
# Create an ordered factor for 'month'
order_of_months <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
Airbnb_Texas$month <- factor(Airbnb_Texas$month, levels = order_of_months)

# Plot
ggplot(Airbnb_Texas %>% 
         group_by(month) %>%
         summarize(mean_rate = mean(average_rate_per_night, na.rm = TRUE)),
       aes(month, mean_rate, fill = month, color = month)) +
  geom_col() +
  labs(title = "Average Rate Per Month") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

January has the highest average rate per night.

## 6- Average rate per night per year

```{r}
ggplot(Airbnb_Texas %>% 
         group_by(year) %>%
         summarize(mean_rate = mean(average_rate_per_night, na.rm = TRUE)),
       aes(x = year, y = mean_rate, group = 1)) +
  geom_line() +
  labs(title = "Average Rate Over Years")

```

Over the years, the average rate per night grew due to increased demand and the cost of life. Rents are different from 2008 to 2017.

## 7- Number of bedrooms 

One bedroom is the most popular.

```{r}
Airbnb_Texas$bedrooms_count[Airbnb_Texas$bedrooms_count=="Studio"] <- "1"
Airbnb_Texas$bedrooms_count <- as.numeric(Airbnb_Texas$bedrooms_count)
 ggplot(Airbnb_Texas , aes(x = bedrooms_count, fill = as.factor(bedrooms_count))) +
  geom_histogram(binwidth = 0.5, position = "identity", alpha = 0.7) +
  scale_fill_manual(values = rainbow(length(unique(Airbnb_Texas$bedrooms_count)))) +
   labs(title = "Number of Bedrooms")

```

## 8-Average rate night per bedroom

```{r}
Airbnb_Texas %>% 
  group_by(bedrooms_count) %>%
  summarize(mean_rate = mean(average_rate_per_night, na.rm = TRUE))
```

## Conclusion
From 2008 to 2017, 18259 were listed in Texas, but massive growth started around 2011. When we look at the top 10 cities with the most listings, we realize they are the big cities of Texas with many visitors, international airports, attractions... The average rate per night has considerably grown over the years. One bedroom is the most popular listing because it's the cheapest average night. Overall, Airbnb is a trendy business in Texas, and the questions we answered should help investors direct themselves into this business.
